\chapter{Ship Design}
AAUSHIP1 has been designed by the ASV group during their 6th semester and is extensively designed in a 3D modeling program called RhinoCeros. The plan is to get the ship molded in plastic. As this process is expensive, it's quite important that we agree on the final design of the ship before we send it to the processing company. 

This ship is designed using a tool called Lofting, which takes some lines that runs along the ship, and then generates a surface from these. As none of us have any experience in creating ships - the design is loosely based on what we "think" is a good hull design. 

However, to further strengthen the structure, inlays are made by slicing a box at the edges of the ship hull. These are then to be produced in the machine workshop at the department of electronics. 

To get a better understanding of ship dynamics a rapid prototyping technique have been used, where 3D models of the ship have been printed - and if it didn't live up to our demands (visually deciding if the design was flawed), a revised model was then printed. Using an iterative process - the final design was reached. 

\section{Parts list}
To develop the ship, a wide range of components had to be chosen. There are several things to be taken into consideration when designing a scale ship from scratch. The first component decided upon is the engines. As the ship should be able to cope with rather powerfull streams a powerfull propulsion system had to be chosen. The main propulsion unit consists of 2 Grapuner Brushless 750 14.8 V 1200W engines - which together produces around 3 HP at maximum input. With these powerfull engines we have a large dynamic range, that allows for fast acceleration and a fast maneuvering, as well as navigating in strong currents. 

The propellers are 2 counterrotating Raboesch brass propellers with a radius of 5 cm. 

For power, the ship is fitted with 6 4-celled Lithium-Polymer batteries, each holding up to 3.2 Ah. These will be divided into different sections, 4 serving the main engines, and 2 serving the onboard controls. If the batteries run at full power (draining a maximum of 80 Amps) the ship will be able to sail for 5 minutes, however, a scenario where the engine runs at full power with a brake on the propellers are unlikely.

Measurements show that the engines run at 60/60 (with a 500 max) they drain 2.5As combined, this will give an operation time of 5.1 hours - thus giving quite a big range on the ships. Assuming a 1 m/s velocity kept constantly for 2.5 hours, this gives a range of 9 kilometers forth and 9 kilometers back or being able to measure an area of 1 x 1 km in strips 100 meters apart and still have surplus power.